---
title: "ESM244 Lab1"
author: "Hannah Garcia-Wickstrum"
date: "1/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(sf)
library(tmap)
library(dplyr)
```

### Read in SF trees data

```{r}
sf_trees <- read_csv(here::here("data", "sf_trees", "sf_trees.csv"))
```

### Basic wrangling reminders

Refresh some skills for data wrangling & summary statistics using functions in the 'dplyr' package. 

Find the top 5 highest observations of trees by legal status, do some wrangling, and make a graph.

```{r}
top_5_status <- sf_trees %>% 
  count(legal_status) %>% #count replaces group_by and summarize_n and puts it into a table
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  relocate(tree_count) %>%  
  slice_max(tree_count, n = 5)
```

Make a graph of those top 5 observations by legal status.
```{r}
ggplot(data = top_5_status, aes(x= fct_reorder(legal_status, tree_count), y= tree_count)) +
  geom_col() + # default is to list by alphabetical order 
  labs(x= "Legal Status", y= "Tree Count") +
  coord_flip() + # flips the axes
  theme_minimal()
```

### A few more data wrangling refresher examples

Only want to keep observations (rows) for Blackwood Acacia Trees.

```{r}
blackwood_acacia <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>% # good for not exact matches but if you're looking to see if it contains any keywords of interest
  select(legal_status, date, latitude, longitude)

ggplot(data=blackwood_acacia, aes(x=longitude, y=latitude)) +
  geom_point()
```

### tidyr::separate() and unite() functions

Useful for combining and separating columns

```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = "::") # column you want to split, give names of two new columns, and how it should separate the text
```

Example: tidyr::unite()

```{r}
sf_trees_unite <- sf_trees %>% 
  unite("id_status", tree_id:legal_status, sep ="_cool!_") # name of new column, columns you want to combine, and how you want to separate them
```

### Make some actual maps of blackwood acacia trees in SF

'st_as_sf()' to convert latitude and longitude ti spatial coordinates.

```{r}
blackwood_acacia_sp <- blackwood_acacia %>% 
  drop_na(longitude, latitude) %>% 
  st_as_sf(coords = c("longitude", "latitude")) 

# need to assign a reference coordinate system
st_crs(blackwood_acacia_sp) = 4326

ggplot(data=blackwood_acacia_sp) +
  geom_sf(color="darkgreen")
```

Read in SF roads shapefile:
```{r}
sf_map <- read_sf(here("data", "sf_map", "tl_2017_06075_roads.shp"))

st_transform(sf_map, 4326)

ggplot(data=sf_map) +
  geom_sf()
```

Combine blackwood acacia tree observations & SF roads map:
```{r}
ggplot() +
  geom_sf(data=sf_map, size = 0.1, color = "darkgray") +
  geom_sf(data=blackwood_acacia_sp, color = "red", size = 0.5) +
  theme_void()
```

Now to create an interactive map:
```{r}
tmap_mode("view") #default is a static map

tm_shape(blackwood_acacia_sp) +
  tm_dots()
```










